\begin{abstract}
Conformal prediction provides rigorous, distribution-free uncertainty guarantees, but often yields prohibitively large prediction sets in structured domains such as routing, planning, or sequential recommendation. We introduce \emph{graph-based conformal compression}, a framework for constructing compact subgraphs that preserve statistical validity while reducing structural complexity. We formulate compression as selecting a smallest subgraph capturing a prescribed fraction of the probability mass, and reduce to a weighted version of densest $k$-subgraphs in hypergraphs, in the regime where the subgraph has a large fraction of edges. We design efficient approximation algorithms that achieve constant factor coverage and size trade-offs. Crucially, we prove that our relaxation satisfies a \emph{monotonicity} property, derived from a connection to parametric minimum cuts, which guarantees the nestedness required for valid conformal guarantees. 
%Our results therefore not only highlight an algorithmic regime, distinct from classical densest-$k$-subgraph hardness settings, where the problem can be approximated efficiently, but also bridge efficient conformal prediction with combinatorial graph compression via monotonicity. 
 Our results on the one hand bridge efficient conformal prediction with combinatorial graph compression via monotonicity, to provide rigorous guarantees on both statistical validity, and compression or size. On the other hand, they also highlight an algorithmic regime, distinct from classical densest-$k$-subgraph hardness settings, where the problem can be approximated efficiently. 
%
We finally validate our algorithmic approach via simulations for trip planning and navigation, and compare to natural baselines.
%



%Conformal prediction provides rigorous uncertainty guarantees for model outputs but can produce prohibitively large prediction sets in structured domains such as routing, planning, or sequential recommendation. We introduce \emph{graph-based conformal compression}, a framework for constructing compact subgraphs that preserve the statistical validity of conformal prediction while reducing structural complexity. We study a formulation that selects a smallest subgraph capturing a prescribed fraction of the probability mass,  and reduce to a weighted version of densest $k$-subgraphs in hypergraphs, in the regime where the subgraph has a large fraction of edges. We design efficient approximation algorithms that achieve constant factor coverage and size trade-offs.   Our results highlight an algorithmic regime, distinct from classical densest-$k$-subgraph hardness settings, where the problem can be approximated efficiently, bridging conformal prediction with combinatorial graph compression. We finally validate our algorithmic approach on synthetic and real-world instances of trip planning and navigation, showing in each case that our approach handily beats natural baselines.
\end{abstract}